### 2016-03-30 main.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Masters/Doctoral Thesis 
% LaTeX Template
% Version 2.1 (2/9/15)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Version 2.0 major modifications by:
% Vel (vel@latextemplates.com)
%
% Original authors:
% Steven Gunn  (http://users.ecs.soton.ac.uk/srg/softwaretools/document/templates/)
% Sunil Patel (http://www.sunilpatel.co.uk/thesis-template/)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%------------------------------------------------------------------------------

\documentclass[
11pt, % The default document font size, options: 10pt, 11pt, 12pt
%oneside, % Two side (alternating margins) for binding by default, uncomment to switch to one side
english, % ngerman for German
onehalfspacing %singlespacing, % Single line spacing, alternatives: onehalfspacing or doublespacing
%draft, % Uncomment to enable draft mode (no pictures, no links, overfull hboxes indicated)
%nolistspacing, % If the document is onehalfspacing or doublespacing, uncomment this to set spacing in lists to single
%liststotoc, % Uncomment to add the list of figures/tables/etc to the table of contents
%toctotoc, % Uncomment to add the main table of contents to the table of contents
%parskip, % Uncomment to add space between paragraphs
]{MastersDoctoralThesis} % The class file specifying the document structure

\usepackage[utf8]{inputenc} % Required for inputting international characters
\usepackage[T1]{fontenc} % Output font encoding for international characters

\usepackage{palatino} % Use the Palatino font by default

%\usepackage[backend=bibtex,style=trad-unsrt,sorting=nyt,natbib=true,isbn=false,doi=false,url=false]{biblatex} % User the bibtex backend with the authoryear citation style (which resembles APA)
%\addbibresource{C2_preliminaries.bib} % The filename of the bibliography
%\addbibresource{C3_turbulence.bib} % The filename of the bibliography
%\addbibresource{C4_RB_VMS.bib} % The filename of the bibliography
%\addbibresource{C5_TBT_OSS.bib} % The filename of the bibliography
%\addbibresource{C6_SRK.bib} % The filename of the bibliography
%\addbibresource{C7_SVMS.bib} % The filename of the bibliography
%\addbibresource{C8_NACA.bib} % The filename of the bibliography
\bibliographystyle{abbrv}

\usepackage[]{csquotes} % Required to generate language-dependent quotes in the bibliography

\usepackage{amssymb,amsmath}
\usepackage{amsfonts}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{pifont}
\usepackage{amsthm}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{multicol}
\usepackage[ruled,vlined]{algorithm2e}

%------------------------------------------------------------------------------
%	INCLUDE DEFINITIONS FILE
%------------------------------------------------------------------------------
\include{definitions}

%------------------------------------------------------------------------------
%	THESIS INFORMATION
%------------------------------------------------------------------------------

\thesistitle{Gradient Jump Viscosity Shock Capturing Techniques} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
\supervisor{Prof. Santiago \textsc{Badia}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
\examiner{} % Your examiner's name, this is not currently used anywhere in the template, print it elsewhere with \examname
\degree{Doctor of Philosophy} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
\author{Alba \textsc{Hierro Fabregat}} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
\addresses{} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname

\subject{} % Your subject area, this is not currently used anywhere in the template, print it elsewhere with \subjectname
\keywords{} % Keywords for your thesis, this is not currently used anywhere in the template, print it elsewhere with \keywordnames
\university{\href{http://www.upc.edu}{Universitat Politècnica de Catalunya}} % Your university's name and URL, this is used in the title page and abstract, print it elsewhere with \univname
\program{Doctorat en Matemàtica Aplicada}
\department{Departament d'Enginyeria Civil i Ambiental} % Your department's name and URL, this is used in the title page and abstract, print it elsewhere with \deptname
\group{} %{\href{http://www.cimne.com/spacehome/2/1156}{Large Scale Scientific Computing group}} % Your research group's name and URL, this is used in the title page, print it elsewhere with \groupname
\faculty{\href{https://www.fme.upc.edu}{Facultat de Matemàtiques i Estadística}} % Your faculty's name and URL, this is used in the title page and abstract, print it elsewhere with \facname

\hypersetup{pdftitle=\ttitle} % Set the PDF's title to your title
\hypersetup{pdfauthor=\authorname} % Set the PDF's author to your name
\hypersetup{pdfkeywords=\keywordnames} % Set the PDF's keywords to your keywords

\begin{document}

\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the pre-content pages

\pagestyle{plain} % Default to the plain heading style until the thesis style is called for the body content

%------------------------------------------------------------------------------
%	TITLE PAGE
%------------------------------------------------------------------------------

\begin{titlepage}
\begin{center}

\textsc{\LARGE \univname}\\[1.5cm] % University name
\textsc{\Large Doctoral Thesis}\\[0.5cm] % Thesis type

\HRule \\[0.4cm] % Horizontal line
{\huge \bfseries \ttitle}\\[0.4cm] % Thesis title
\HRule \\[1.5cm] % Horizontal line

\begin{minipage}{0.4\textwidth}
\begin{flushleft} \large
\emph{Author:}\\
%\href{http://www.johnsmith.com}{\authorname} % Author name - remove the \href bracket to remove the link
{\authorname} % Author name - remove the \href bracket to remove the link
\end{flushleft}
\end{minipage}
\begin{minipage}{0.4\textwidth}
\begin{flushright} \large
\emph{Supervisor:} \\
%\href{http://www.jamessmith.com}{\supname} % Supervisor name - remove the \href bracket to remove the link  
{\supname} % Supervisor name - remove the \href bracket to remove the link  
\end{flushright}
\end{minipage}\\[3cm]

\large \textit{A thesis submitted in fulfilment of the requirements\\ for the degree of \degreename}\\[0.3cm] % University requirement text
\textit{in the}\\[0.4cm]
%\groupname\\\deptname\\[2cm] % Research group name and department name
\progname\\[0.5cm]
\deptname\\[2cm] % Research group name and department name

{\large Barcelona, February 2016}\\[2cm] % Date+
\includegraphics[width=8cm]{../Figures/Cover/logo_fme.png}
%\includegraphics{Logo} % University/department logo - uncomment to place it

\vfill
\end{center}
\end{titlepage}

%------------------------------------------------------------------------------
%	DEDICATION
%------------------------------------------------------------------------------

\dedicatory{To \ldots}

%------------------------------------------------------------------------------
%	DECLARATION PAGE
%------------------------------------------------------------------------------

%\begin{declaration}
%\addchaptertocentry{\authorshipname}
%
%\noindent I, \authorname, declare that this thesis titled, \enquote{\ttitle} and the work presented in it are my own. I confirm that:
%
%\begin{itemize} 
%\item This work was done wholly or mainly while in candidature for a research degree at this University.
%\item Where any part of this thesis has previously been submitted for a degree or any other qualification at this University or any other institution, this has been clearly stated.
%\item Where I have consulted the published work of others, this is always clearly attributed.
%\item Where I have quoted from the work of others, the source is always given. With the exception of such quotations, this thesis is entirely my own work.
%\item I have acknowledged all main sources of help.
%\item Where the thesis is based on work done by myself jointly with others, I have made clear exactly what was done by others and what I have contributed myself.\\
%\end{itemize}
% 
%\noindent Signed:\\
%\rule[0.5em]{25em}{0.5pt} % This prints a line for the signature
% 
%\noindent Date:\\
%\rule[0.5em]{25em}{0.5pt} % This prints a line to write the date
%\end{declaration}
%
\cleardoublepage

%------------------------------------------------------------------------------
%	QUOTATION PAGE
%------------------------------------------------------------------------------

\vspace*{0.2\textheight}

%\noindent\enquote{\itshape Thanks to my solid academic training, today I can write hundreds of words on virtually any topic without possessing a shred of information, which is how I got a good job in journalism.}\bigbreak
%
%\hfill Dave Barry

\noindent\enquote{\itshape La teoria sense pràctica és inútil
Però la pràctica sense teoria, és efímera.}\bigbreak

\hfill Brigitte Vasallo

%------------------------------------------------------------------------------
%	ABSTRACT PAGE
%------------------------------------------------------------------------------

\begin{abstract}
\addchaptertocentry{\abstractname} % Add the abstract to the table of contents

%The Thesis Abstract is written here (and usually kept to just this page). The page is kept centered vertically so can expand into the blank space above the title too\ldots
%\input{abstract}

\end{abstract}

%------------------------------------------------------------------------------
%	RESUM
%------------------------------------------------------------------------------

\begin{resum}
\addchaptertocentry{\resumname} % Add the abstract to the table of contents

%\input{resum}

\end{resum}

%------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%------------------------------------------------------------------------------

\begin{acknowledgements}
\addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
Dinar del mes
Filipos
JIPI

%The acknowledgements and the people to thank go here, don't forget to include your project advisor\ldots
%\input{acknowledgements}

\end{acknowledgements}

%------------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES PAGES
%------------------------------------------------------------------------------

\tableofcontents % Prints the main table of contents

\listoffigures % Prints the list of figures

\listoftables % Prints the list of tables

%------------------------------------------------------------------------------
%	ABBREVIATIONS
%------------------------------------------------------------------------------

\begin{abbreviations}%{ll} % Include a list of abbreviations (a table of two columns)

\textbf{DMP} & \textbf{D}iscrete \textbf{M}aximum \textbf{P}rinciple \\

\end{abbreviations}

%------------------------------------------------------------------------------
%	PHYSICAL CONSTANTS/OTHER DEFINITIONS
%------------------------------------------------------------------------------

%\begin{constants}%{lr@{${}={}$}l} % The list of physical constants is a three column table
%
% %The \SI{}{} command is provided by the siunitx package, see its documentation for instructions on how to use it
%
%Speed of Light & $c$ & \SI{2.99792458e8}{\meter\per\second} (exact)\\
%%Constant Name & $Symbol$ & $Constant Value$ with units\\
%
%\end{constants}

%------------------------------------------------------------------------------
%	SYMBOLS
%------------------------------------------------------------------------------

%\begin{symbols}%{lll} % Include a list of Symbols (a three column table)
%\input{symbols_definition}
%\end{symbols}

%------------------------------------------------------------------------------
%	THESIS CONTENT - CHAPTERS
%------------------------------------------------------------------------------

\mainmatter % Begin numeric (1,2,3...) page numbering

\pagestyle{thesis} % Return the page headers back to the "thesis" style

% Include the chapters of the thesis as separate files from the Chapters folder
% Uncomment the lines as you write the chapters

\input{Chapters/paper1}

\input{Chapters/paper2}

%------------------------------------------------------------------------------
%	THESIS CONTENT - APPENDICES
%------------------------------------------------------------------------------

\appendix % Cue to tell LaTeX that the following "chapters" are Appendices

% Include the appendices of the thesis as separate files from the Appendices folder
% Uncomment the lines as you write the Appendices

%\input{Appendices/AppendixA}

%------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%------------------------------------------------------------------------------

%\printbibliography[heading=bibintoc]
\bibliography{art004,ar009,hp-adaptivity}

%------------------------------------------------------------------------------

\end{document}