### 2016-03-31 symbols_definition.tex
minor changes

\begin{symbols}

\multicolumn{3}{c}{{\Large \textbf{Lower-case Roman}}}\\
$d$ & space dimension\\
$n$ & unit outside normal\\
$u(\x,t)$      & variable of interest       &  \\

\addlinespace % Gap to separate the Roman lower from Roman Capital
\addlinespace

\multicolumn{3}{c}{{\Large \textbf{Upper-case Roman}}}\\
\addlinespace % Gap to separate the Roman lower from Roman Capital
\addlinespace

\multicolumn{3}{c}{{\Large \textbf{Lower-case Greek}}}\\
$\beta$ & convection term\\
$ \tau_c $ & continuity stabilization parameter & \\

\addlinespace % Gap to separate the Greek lower symbols from the Greek Capital
\addlinespace

\multicolumn{3}{c}{{\Large \textbf{Upper-case Greek}}}\\
$\Omega$ & domain \\
$\Gamma$  & domain boundary ($\Gamma=\partial\Omega$) & \\
$\Gamma_{\rm in}$ & inflow boundary ($\Gamma_{\rm in} := \{ x \in \partial \Omega: \, \beta(x) \cdot n(x) < 0\}$)\\
\addlinespace % Gap to separate the Greek capital symbols from the subscripts
\addlinespace

\multicolumn{3}{c}{{\Large \textbf{Superscripts and subscripts}}}\\
$ (\cdot)_h $ & Finite Element component & \\
$ (\cdot)_{n+1} $ & current time step & \\
$ (\cdot)_n $ & previous time step & \\

\end{symbols}