### 2016-04-01 symbols_definition.tex
few changes

\begin{symbols}

\multicolumn{3}{c}{{\Large \textbf{Lower-case Roman}}}\\
$d$ & space dimension& \\
$n$ & unit outside normal& \\
$u(\x,t)$      & variable of interest       &  \\

\addlinespace % Gap to separate the Roman lower from Roman Capital
\addlinespace

\multicolumn{3}{c}{{\Large \textbf{Upper-case Roman}}} \\
$\mathcal{N}_h(K)$ & Set of interpolation nodes & \\	
$\vh$ & piecewise polynomial space of discrete solutions (continuous or discontinuous)& \\
$V_h^K$ & piecewise polynomial space of order $k$  &\\
$V_{h,0}$ & piecewise polynomial space with homogeneous Dirichlet boundary conditions  &\\
$\vm$ & piecewise polynomial space of discrete solutions at the $m$ adaptive step & \\

\addlinespace % Gap to separate the Roman lower from Roman Capital
\addlinespace

\multicolumn{3}{c}{{\Large \textbf{Lower-case Greek}}} \\
$\beta$ & convection term& \\
$ \tau_c $ & continuity stabilization parameter & \\

\addlinespace % Gap to separate the Greek lower symbols from the Greek Capital
\addlinespace

\multicolumn{3}{c}{{\Large \textbf{Upper-case Greek}}}\\
$\Omega$ & domain \\
$\Gamma$  & domain boundary ($\Gamma=\partial\Omega$) & \\
$\Gamma_{\rm in}$ & inflow boundary ($\Gamma_{\rm in} := \{ x \in \partial \Omega: \, \beta(x) \cdot n(x) < 0\}$)\\
\addlinespace % Gap to separate the Greek lower symbols from the Greek Capital
\addlinespace

\multicolumn{3}{c}{{\Large \textbf{Upper-case Greek}}}\\
$\jump{\cdot}$ & Jump operator & \\
$\mean{\cdot}$ & Average operator & \\
\addlinespace % Gap to separate the Greek capital symbols from the subscripts
\addlinespace

\multicolumn{3}{c}{{\Large \textbf{Superscripts and subscripts}}}\\
$ (\cdot)_h $ & Finite Element component & \\
$ (\cdot)_{n+1} $ & current time step & \\
$ (\cdot)_n $ & previous time step & \\

\end{symbols}