### 2016-04-12 Chapters/paper4.tex
added new paper

\chapter{Monotonicity preserving methods for discontinuous Galerkin Methods}

We define a \textit{nonnegative type} matrix as any $N\times N$ matrix $A=\{a_{ij}\}$ such that
\begin{align}
&a_{ij} \leq 0  & & {\rm for}\, i\neq j, \, i=1,\cdots,N_R, j=1,\cdots,N \\
&\sum_{j=1}^{N} a_{ij} \geq 0 & &{\rm for}\, i=1,\cdots,N_R
\end{align}
We define $A_R = \{a_{ij}\}$ with $i,j=1,\cdots, N_R$.
In \cite{codina_discontinuity-capturing_1993}, Codina proved that
\begin{theorem}[Sufficient condition for DMP]
	If A is of \textit{non-negative type}, $A_R$ is non singular and $b_i\leq 0$ for $i=1,\cdots,N_r$ then 
	\begin{align*}
	Ax=b & \Rightarrow & \exists m\in \{N_R+1,\cdots,N\} \, s.t. \, \max_{i\in\{1,\cdots,N\}}\{x_i\}=x_m
	\end{align*}
\end{theorem}

Somehow this is changed to the column property and dG methods already have such property.