### 2016-06-06 symbols_definition.tex

\begin{symbols}

\multicolumn{3}{c}{{\Large \textbf{Lower-case Roman}}}\\
$d$ & space dimension& \\
$n$ & unit outside normal& \\
$u(\x,t)$      & variable of interest       &  \\

\addlinespace % Gap to separate the Roman lower from Roman Capital
\addlinespace

\multicolumn{3}{c}{{\Large \textbf{Upper-case Roman}}} \\
$\facet$ & facet & \\
$H^\beta(\Omega)$ & Space of solutions of the weak transport problem &\\
& { ($H^\beta(\Omega) := \{ u \in L^2(\Omega): \,  \nabla \cdot (\beta u) \in L^2(\Omega) \}$)}\\
$K$, $K_i$ & elements of the mesh (in Chaper \ref{chap-paper3} they are ordered by convenience of the analysis) & \\
$K^+_\facet$/$K^-_\facet$ & elements that surround facet $\facet$ ($\facet = \partial K^+_\facet\cap K^-_\facet$ ) & \\
$\mathcal{N}_h(K)$ & Set of interpolation nodes & \\
$P_k(\omega)$ & space of complete polynomials of degree $k$ on $\omega$ & \\
$Q_k(\omega)$ & space of polynomials with maximum degree $k$ in $\omega$\\
$\thn$ & finite element mesh & \\
$\thm$ & finite element adaptive mesh at adaptivity step $m$ &\\	
$\vh$ & piecewise polynomial space of discrete solutions (continuous or discontinuous)& \\
$V_h^k$ & piecewise polynomial space of order $k$  &\\
$V_{h,0}$ & piecewise polynomial space with homogeneous Dirichlet boundary conditions  &\\
$\vm$ & piecewise polynomial space of discrete solutions at the $m$ adaptive step & \\
$W^{l,p}$ & Sobolev space & \\
\addlinespace % Gap to separate the Roman lower from Roman Capital
\addlinespace

\multicolumn{3}{c}{{\Large \textbf{Lower-case Greek}}} \\
$\beta$ & convection term& \\
$\kappa_h$ & Scott-Zhang projector & \\
$\varsigma$ & node-to-element map & \\
$ \tau_c $ & continuity stabilization parameter & \\

\addlinespace % Gap to separate the Greek lower symbols from the Greek Capital
\addlinespace

\multicolumn{3}{c}{{\Large \textbf{Upper-case Greek}}}\\
$\Omega$ & domain \\
$\Omega_i$ & macroelement associated to node $x_i$ & \\
$\Gamma$  & domain boundary ($\Gamma=\partial\Omega$) & \\
$\Gamma_{\rm in}$ & inflow boundary ($\Gamma_{\rm in} := \{ x \in \partial \Omega: \, \beta(x) \cdot n(x) < 0\}$)\\

\addlinespace % Gap to separate the Greek capital symbols from the subscripts
\addlinespace

\multicolumn{3}{c}{{\Large \textbf{Superscripts and subscripts}}}\\
$ (\cdot)_h $ & Finite Element component & \\
$ (\cdot)_{n+1} $ & current time step & \\
$ (\cdot)_n $ & previous time step & \\
\addlinespace % Gap to separate the Greek lower symbols from the Greek Capital
\addlinespace

\multicolumn{3}{c}{{\Large \textbf{Other symbols}}}\\
$\jump{\cdot}$ & Jump operator & \\
$\mean{\cdot}$ & Average operator & \\
$\|\cdot\|$ & $L^2(\Omega)$ norm & \\
$\|\cdot\|_\omega$ & $L^2(\omega)$ norm for $\omega\in\Omega$ & \\
\end{symbols}